---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We will first initialize all the packages that we will need to be using.
Then we can download and load in the data from the provided URL, as well as
unzip it and have it ready for use. Before moving onto the next step, we will
finish this section off by reading the csv data from the url file to our
program. 

```{r}
library(knitr)
library(data.table)
library(ggplot2)

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip')
              , method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")

activityDT <- data.table::fread(input = "data/activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
Total_Steps <- activityDT[, c(lapply(.SD, sum, na.rm = FALSE))
                          , .SDcols = c("steps"), by = .(date)] 
head(Total_Steps, 10)

ggplot(Total_Steps, aes(x = steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Steps", y = "Frequency")

Total_Steps[, .(Mean_Steps = mean(steps, na.rm = TRUE)
                , Median_Steps = median(steps, na.rm = TRUE))]
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
